Who Cares About Crackdowns? Exploring the Role of Trust in Individual Philanthropy

### 2021-04-17 manuscript/manuscript.Rmd
A bunch of edits

In 2016, Human Rights Watch claimed that civil society was under more aggressive attack than at any time in recent memory [@Roth:2016]. Governments across the globe have cracked down on civil society organizations (CSOs) and non-governmental organizations (NGOs)—a phenomenon known as "closing civic space" [@CarothersBrechenmacher:2014; @CIVICUS:2017; @DPR:2015; @DPR:2016].[^NGOdef] Legal restrictions, or what we refer to as legal crackdowns, are a core part of these efforts. These crackdowns create barriers to entry, funding, and advocacy for NGOs in an effort to control, obstruct, and repress these organizations. Barriers to funding are especially pervasive and restrict the ability of NGOs to secure financial resources. Restrictive states may prevent the transfer of foreign funds to NGOs based on the origin of these funds and which issues these funds can be used towards.

As a result of these laws, foreign aid (also referred to as public or official aid) channeled through NGOs has decreased considerably in repressive countries [@Brechenmacher:2017; @ChaudhryHeiss:2018; @DupuyPrakash:2018]. However, philanthropy from private donors and foundations is not as adversely affected [@McGill:2018]. Many foundations have continued channeling funds to countries with hostile legal environments [@NeedlesWolcheckSato:2018]. Given this trend of reduced public aid following restricted legal environments for NGOs abroad, INGOs may need to devote more resources to fundraise from individual donors.

Individual private donations are a critical source of INGO funding. While total private philanthropy[^phildef] in the United States has grown substantially over time, it is notable that private giving to NGOs working in international affairs  and international affairs-focused organizations are one of only two categories of organizations that have continued experiencing substantial growth in giving and reached \$22.08 billion in 2018 [@givingusa:2019].[^intlaffairs] Most of this growth is driven by individuals making small-scale or micro-donations.[^smalldonors] In 2017, 70% of donations to international affairs NGOs came from individuals, rising to 86% in 2018 [@givingusa:2019]. But relying on individual donors means that NGOs must consider their preferences when fundraising. Which individual donor characteristics and attributes are most salient for INGOs interested in maximizing philanthropy at a time when they are facing increasingly hostile environments abroad?

There exists a robust scholarship on the determinants of individual-level philanthropy to domestic causes. But the dynamics of *international* giving to NGOs are under-examined in the literature on individual donor attributes and preferences.[^exceptions] Motivating individuals to donate to international causes is hard as the number of recipients are large and far removed from the donor [@CasaleBaumann:2015, 100]. Trust is an essential component for INGO sustainability and for many nonprofits, trust often translates into grants and donations [@coombs2007]. Geography often impedes donors' direct evaluation of their work and trust is one of the most influential criteria that private donors use when deciding which nonprofits to support [@prakash2010trust; @bekkers2003trust]. A lack of trust could be further fueled by legal trouble with the NGO host government—a dynamic that individuals may not wish to enter with their dollars. How does trust in NGOs influence individuals' preferences to donate, especially when groups face crackdown?

We explore this using a simulated market for philanthropic donations towards INGOs, based on data from a nationally representative sample of individuals in the United States who regularly donate to charity. We adapt methods used in marketing to generate dozens of distinct simulated personas that represent typical donor profiles, and we vary these personas along different dimensions of demographic, political, and social attributes. We find that an organization's target issue area and its relationship with its host government interact strongly with individuals' characteristics. Personas with high levels of trust in political and charitable institutions and who are actively engaged in charity and volunteerism prefer donating to human rights organizations that have friendly relationships with their host governments. In contrast, personas with low levels of social trust eschew both human rights organizations *and* any type of NGO with more contentious relationships with host governments. Instead, they prefer donating to NGOs with friendly host-country relationships and working on less contentious issues like emergency response and refugee relief. Personas with higher trust are more comfortable donating to legally besieged INGOs. Our findings imply that individual donors with high trust would continue remaining supportive of INGOs when they face criticism and crackdown from foreign governments. On the other hand, government crackdown may further sour NGOs least trusting individual donors.

This study contributes both to scholarship and practice in public policy, nonprofits, and philanthropy in three important ways. It helps us understand how the international community and NGOs might respond to the phenomenon of closing civic space, which does not appear to be receding. Most NGOs working internationally have previously relied on government and foundation funding. But in the face of decreasing funds from official donors, our results show that INGOs can rely on individual donors, specifically, their most trusting donors, when groups face a hostile environment abroad. These findings lend support to previous literature arguing that trust matters  when individuals donate to organizations acting under great uncertainty or working towards difficult goals [@Wiepking:2010].

Second, knowledge of demographic traits and its impact on funding can help nonprofits craft more effective and targeted marketing strategies [@Rajan:2009, 414]. Our paper shows that in readjusting their fundraising strategies towards individuals, framing, or the process through which actors present information to influence perceptions, is particularly important [@chong2007framing]. Framing affects their willingness to donate. However, existing research is unclear about which frames are effective when making appeals to individual donors considering international philanthropy. Since our study looks at multiple organizational and host country factors, our results can help NGOs frame their appeals. In particular, our results show that NGOs may benefit from publicizing crackdown at least towards their most trusting donors.

Finally, while research in public policy examines giving to NGOs, philanthropy is not seen as relevant by international relations as a discipline [@youde2019]. However, philanthropy—particularly donations received by INGOs—can help us understand changes in the dynamics of global governance. Private individuals exercise power and influence in unique ways that differ from other non-state actors. As @youde2019 [p. 44] points out, it changes our conceptions of individuals being passive recipients to active entities that are not merely objects of governance. Thus, it can shed light on the changing logic of global governance.

@@

Foreign aid to civil society—and NGOs in particular—has steadily increased over time. Many government and foundation donors perceive NGOs as nimbler and less bureaucratic than governments. In countries with weak institutions and poor governance, where direct aid transfers to governments raise issues of misuse and bureaucratic inefficiency, donors specifically seek out NGOs for projects [@Dietrich2013]. However, NGO-state relationships have evolved over time—in the 1950s and 60s, international assistance channeled through NGOs aimed at producing socio-economic change rather than political progress. But in the aftermath of the dissolution of the Soviet Union, many NGOs had the explicit goal of promoting democracy and bringing down authoritarian governments. Subsequently, many states perceive them as threatening and have sought to repress them.

Mre than 100 countries have proposed or enacted 244 measures restricting, repressing, or shutting down civil society since 2013 (ICNL 2021 CITATION FROM ISA PAPER). This has had a three-hold impact on INGOs. First, it has led to a reduction in resources and grants available to INGOs. Second, restrictive NGO laws can affect INGO programming by preventing organizations from working in certain (contentious) issue areas. To deal with this, some groups have adapted by changing the nature of their programming to appear less threatening to the repressive countries. Finally, in the worst-case scenario, this has led to a loss in INGOs' access to target countries—often also associated with the use of violence against INGOs and their employees. In this paper, we exlusively focus on the first dilemma—the reduction in INGO funding and possible responses to it.

@@

Many INGOs in the Global South receive funds from a variety of public and private sources each year. Public funds—or traditional donor aid—channel money from official aid agencies through a variety of bilateral and multilateral institutions. Private sources include foundations, private corporations, and individuals. While INGOs have not been passive in responding to the crackdown on civil society, they face increasing obstacles from a funding perspective.

Closing civic space not just resulted in net losses of income to INGOs, particularly through official bilateral and multilateral aid [@DupuyPrakash:2018], it has also affected how and where funds are allocated. Following the passage of anti-NGO laws, official donors move funds away from contentious causes such as human rights, media, corruption, advocacy, among others, towards tamer and more regime-compatible causes such as health, education [@ChaudhryHeiss:2018]. The latter trend has also been exacerbated by a preference among donors working in official aid and the democracy promotion establishment to gradually increase funding directed towards tame causes, at the expense of more contentious causes. This is because many tame causes can have easily quantifiable, measurable outputs, which makes it clear to show a program's effectiveness over time [@Bush:2015].

We argue that individual private donors can help fill in an important gap. Individual donors do not use the same performance-based metrics as official donors [@desaikharas]. Rather, individual donors often use heuristics to simply their decision-making, making cursory judgments about an organization's issue area, mission, vision, and values and seek out supplementary information from friends, family, and acquaintances [@TBPrakash:2017; @SzperPrakash:2011]. Due to these differences, individual donors can help address some of the challenges arising from reduced public funding available to INGOs. To be clear, we are not arguing that giving by individual donors would be a perfect substitute for NGOs net loss of income or income for certain programs, but rather, an important supplement to reduced funding from foreign sources. Moreover, it would also be a supplement to private funding from foundations towards international nonprofits which has not been as adversely affected, with foundations continue to channel funds to countries with hostile legal environments [@NeedlesWolcheckSato:2018].

However, research on the determinants of individual-level philanthropy for international causes—particularly towards organizations working in challenging environments—is still in its infancy due to the absence of expansive and high-quality data [@greenhill2013]. Moreover, a majority of research on giving to international groups is restricted to elite high net-worth donors[^highnetworth] [@USTrust2014]. We still know little about demographic traits and experiences that motivate or deter individuals to donate during shrinking civic space. Below, we look at key drivers of individual-level philanthropy and show how the importance of trust in donating during civic crackdowns.

@@

Social capital, particularly associational capital, or "the networks of formal and informal relationships to which people are associated," is often considered an important indicator of giving [@havens2007, 240]. Volunteering exposes individuals to the need for donations, making them feel more compelled to donate [@hossainlamb2012]. Those who participate in a variety of associations are both more likely to donate as well as make larger donations [@hossainlamb2017].

Higher education also influences charitable-giving decisions as it fosters prosocial motivations. Education brings people into social networks that entail a higher level of solicitation, which is vital in giving to international causes [@BekkersWiepking:2011]. Education also helps determine the kinds of causes that people donate to globally. @Micklewright:2009 show that higher education matters more for donating to international relief organizations than donating to domestic organizations.

A number of NGOs facing crackdowns abroad also work on issues that may be seen as "contentious," including human rights, elections, corruption, advocacy, and media freedom—issues that may have a harder time attracting funds due to their non-essential nature when compared to causes like health, education, sanitation, and social services [@Bush:2015]. However, additional education has been shown to increase abstract thinking [@wiepkingmaas, 1978], which may be necessary to donate to causes that do not have quantifiable goals, including politically sensitive or contentious causes mentioned above.

@@

We explore this question using a conjoint experiment conducted with a nationally representative sample in the U.S. Importantly, because there is so little research about international donations during legal crackdown by host governments, our analysis and results are exploratory and descriptive. Accordingly, we do not test specific hypotheses or theories.

TODO: Fix final point - connect testing stuff with existing theories - we don't have a list of things to test

However, we do talk about causal effects given that we conduct an experiment that varies a number of factors in order to elicit opinions from respondents. Instead, we structure our analysis around broader themes from the existing literature and investigate how donors' (1) demographic characteristics, (2) public affairs knowledge and experience, and (3) levels of trust in government, charitable work, and associational life change individual donation preferences across different organizational issue areas, relationships with host governments, and funding sources. Our innovative experimental approach provides rich quantitative details of individual and organizational characteristics simultaneously and reveals unique insights into donor motivations that both confirm and expand previous research.

@@

We use a conjoint experiment to study the simultaneous effects of organizational characteristics and individual attributes on individuals' preferences to donate to a nonprofit organization. Conjoint analysis allows us to elicit donor preferences without respondents explicitly stating their exact preferences. This kind of analysis is commonly used in marketing to reveal consumer preferences for hypothetical products, and has become increasingly popular in political science and public policy [@Helfer:2016; @Mummolo:2016; @KnudsenJohannesson:2019]. Because we are interested in exploring the complexity behind the decision to donate, we apply conjoint methods to the study of nonprofits and philanthropy.

In our experiment, we presented respondents with four randomly selected international nonprofit organizations with four randomly assigned features.[^appendix-survey] We included organizations that are associated with a range of contentious and noncontentious issues: Amnesty International, Greenpeace, Oxfam, and the International Committee of the Red Cross. Because of their strong name recognition, these organizations (and others like World Vision, the World Wildlife Fund, and Médecins Sans Frontières) are commonly used in experimental surveys on philanthropy [@FaulknerTruongRomaniuk:2015].

@@

```{r organization-attributes-full}
tbl_org_attributes_caption <- "Organization attributes varied in the experiment. Each participant saw repeated combinations of attributes, with one attribute randomly selected from each column."

@@

The hypothetical organizations are based on random combinations of features, including one of four issue areas, the presence or absence of two organizational practices, one of three funding sources, and one of three relationships with host governments (see Table \@ref(tab:organization-attributes-full); see also Q3.1 in the survey text in the appendix). Respondents saw twelve sets of four hypothetical organizations out of 288 (4 × 4 × 2 × 3 × 3) possible combinations, meaning no single respondent was offered every combination. While not every combination necessarily reflected reality—for instance, a respondent might have been asked to compare a fictional Amnesty International that focuses on environmental issues and that is funded by private donations with a Greenpeace that focuses on human rights and that is funded by government grants—it is not important that these features align with true organizational characteristics, since the goal of these repeated hypothetical questions is to identify which attributes are the most salient for donors.

In June 2019, we recruited survey respondents through Centiment, a commercial provider of high-quality, non-probability opt-in survey panels. Centiment ensures panel quality by actively recruiting and paying representative samples of the U.S. population. To participate in the study, respondents were first screened based on their charitable activities. Those who indicated that they gave to charity once every few years or never were disqualified. After screening, we collected a sample of 1,016 respondents, which is sufficient for model estimation.[^sample-size]

@@

We analyze the effect of different combinations of organizational features and individual respondent attributes with a two-level hierarchical Bayesian multinomial logit model.[^model-details] Simply put, the first level of the model predicts individual donor preferences for various combinations of features (e.g. the combination of human rights issues, financial transparency, government funding, and government crackdown), while the second level predicts the heterogeneity of these fitted preferences as a function of individual donor characteristics.

@@

We then combine these different scenarios of donor and organization attributes with the coefficients from a multinomial multilevel regression model to calculate the probability that each simulated persona has of giving to specific combinations of organizational attributes. Any individual characteristics that we do not explicitly vary in Table \@ref(tab:persona-attributes) are held at their means or modal values from the survey sample. Table \@ref(tab:sim-output) provides an example of the simulation output, showing how two different personas are predicted to allocate donations to organizations with various characteristics. The proportion of donations from each persona sums to 100%, as every persona is guaranteed to make a donation.

@@

Relying on simulation methods offers useful analytic benefits. Rather than interpret individual marginal effects from a regression model, we can explore multiple moving parts from the same model simultaneously to understand how individual characteristics interact with each other. This allows us to more fully explore our key research questions and see which kinds of donors are more or less likely to consider specific organizational attributes when deciding to donate, and which donor characteristics are associated with the propensity to donate. Importantly, each persona is an artificial construct. While some survey respondents might match an exact persona profile, others may not. This is acceptable in simulations as each combination of persona characteristics is plausible in the real world.

@@

These simulated findings are more abstract than the results from an ordinary survey sample. But by using simulations, we can discover interesting patterns of donor preferences when faced with organizations focusing on a variety of issues, funding sources, and legal restrictions abroad. Our market simulation reveals which combination of organizational features each persona is most likely to express a preference for donating to. We examine the results of the simulation in three stages, looking at how personas with different (1) demographic characteristics, (2) public affairs knowledge and experience, and (3) levels of trust in government, charitable work, and associational life change their donation preferences across different organizational issue areas, relationships with host governments, and funding sources.

## Patterns across demographics

We begin by looking at differences in donation preferences across a range of basic demographic characteristics. The four panels of Figure \@ref(fig:income) show the average predicted market share for different organizational issue areas, funding sources, and relationships with host governments across different levels of persona income, with two different cutpoints for high and low income. To ease with modeling and interpretation, we split high- and low-income personas based on the 2017 U.S. median income (\$61,372), showing personas with incomes higher and lower than the national median. Because there is wide variation in donor incomes in the U.S., we also show the results from a simulation model with a more extreme split in income, holding annual persona income at \$50,000 and \$100,000+. The difference in the division of high and low income makes no difference in the market shares across issue areas and funding sources, but there are substantial income-based differences across different relationships with host governments.

@@

## Patterns across public affairs knowledge, experience, and social trust

We next look at how differences in public affairs knowledge, experience, and political ideology influence a persona's propensity to donate to an organization. We vary personas along two dimensions: (1) political ideology, with semi-liberal and semi-conservative political leanings (2 and 5 on a 7-point scale), and (2) knowledge and experience with public affairs and the news, with personas that either follow national and international news and that have travelled abroad or personas that do not follow the news or travel.

We combine these dimensions with an additional set of donor characteristics that vary a persona's level of social trust. Personas with high levels of trust have high trust in political institutions and charities, think people should be more charitable, frequently volunteer, donate to charity once a month, have a history of political activism, and are members of an association. Personas with low levels of trust have the opposite characteristics: they do not trust political institutions or charities, rarely donate or volunteer, and are not members of associations.

@@

Personas with high levels of trust, on the other hand, seem to stick with legally besieged organizations and show targeted groups the benefit of the doubt. The global crackdown on NGOs thus sours NGOs' least trusting individual donors. Overall, our results support research that donors with high levels of trust are more active in giving than those with low levels of trust [@sargeant2008] and that trust especially matters when individuals donate to organizations working under great uncertainty and towards difficult goals [@Wiepking:2010].

@@

The conjoint experiment conducted among a nationally representative sample of U.S. respondents with a prior history of charitable giving raises two main concerns for inference. It may be the case that respondent answers—and the corresponding beahavior of simulated personas—could potentially be driven by their priors towards these organizations. In addition, there are also important questions about the generalizability of these results outside the American context. We explore both below.

Given the nature of conjoint experiments, the randomly presented organizations may have influenced respondent choices. All four organizations are well-known, and it is likely that participant responses reflect prior beliefs and news coverage about Amnesty, Oxfam, Greenpeace, and the Red Cross. For instance, in February 2018, news broke out about sexual abuse and misconduct at Oxfam and several other prominent organizations, including Amnesty International and Save the Children. This may have potentially strengthened a narrative that the nonprofit sector suffers from governance failures [@Phillips:2019]. In a competitive NGO environment [@Cooley:2002], maintaining positive perceptions is important as donors who perceive these organizations as untrustworthy may simply take their resources elsewhere. As such, it could be argued that respondents may be less likely to donate to Oxfam or Amnesty because of these scandals.

@@

[^smalldonors]: Micro-donations refer to donations between $0.25–$10; small donors are individuals who donate up to $50 [@SaxtonWang:2014].

@@

[^highnetworth]: High-net worth households have greater than \$200,000 in annual household income and/or have more than \$1,000,000 in assets, excluding primary residence.

>>>>

In 2016, Human Rights Watch claimed that civil society was under more aggressive attack than at any time in recent memory [@Roth:2016]. Governments across the globe have cracked down on civil society organizations (CSOs) and non-governmental organizations (NGOs)—a phenomenon known as "closing civic space" [@CarothersBrechenmacher:2014; @CIVICUS:2017; @DPR:2015; @DPR:2016].[^NGOdef] Legal restrictions, or what we refer to as legal crackdowns, are a core part of these efforts. These crackdowns create barriers to entry, funding, and advocacy for NGOs in an effort to control, obstruct, and repress these organizations. Barriers to funding are especially pervasive and restrict the ability of NGOs to secure financial resources. Restrictive states may prevent the transfer of foreign funds to NGOs based on the origin and purposes of these funds.

As a result of these laws, public or official foreign aid channeled through international NGOs (INGOs) has decreased considerably in repressive countries [@Brechenmacher:2017; @ChaudhryHeiss:2018; @DupuyPrakash:2018]. However, philanthropy from private donors and foundations is not as adversely affected [@McGill:2018]. Many foundations have continued channeling funds to countries with hostile legal environments [@NeedlesWolcheckSato:2018]. Given the reduction of foreign aid following restricted legal environments abroad, INGOs may need to devote more resources to fundraise from individual donors.

Individual private donations are a critical source of INGO funding. While total private philanthropy[^phildef] in the United States has grown over time, it is notable that private giving to NGOs working on international affairs-focused issues are one of only two categories of organizations that have continued experiencing substantial growth in giving, reaching \$22.08 billion in donations in 2018 [@givingusa:2019].[^intlaffairs] Most of this growth is driven by individuals making small-scale (up to \$50) or micro-donations (between \$0.25–\$10) [@SaxtonWang:2014]. In 2017, 70% of donations to international affairs NGOs came from individuals, rising to 86% in 2018 [@givingusa:2019]. But relying on individual donors means that NGOs must consider their preferences when fundraising. Which individual donor characteristics and attributes are most salient for INGOs interested in maximizing philanthropy at a time when they are facing increasingly hostile environments abroad?

A robust scholarship already examines the determinants of individual-level philanthropy to domestic causes. But the dynamics of *international* giving to NGOs are under-examined in the literature on individual donor attributes and preferences.[^exceptions] Motivating individuals to donate to international causes is hard as the number of recipients are large and far removed from the donor [@CasaleBaumann:2015, 100]. Trust is an essential component for INGO sustainability—for many nonprofits, trust often translates into grants and donations [@coombs2007]. Geography often impedes donors' direct evaluation of their work and trust is one of the most influential criteria that private donors use when deciding which nonprofits to support [@prakash2010trust; @bekkers2003trust]. A lack of trust could be further fueled by legal trouble with the INGO host government—a dynamic that individuals may not wish to enter with their dollars. How does trust in INGOs influence individuals' preferences to donate, especially when groups face crackdown?

We explore this using a simulated market for philanthropic donations towards INGOs, based on data from a nationally representative sample of individuals in the United States who regularly donate to charity. We adapt methods used in marketing to generate dozens of distinct simulated personas that represent typical donor profiles, and we vary these personas along different dimensions of demographic, political, and social attributes. We find that an organization's target issue area and its relationship with its host government interact strongly with individuals' characteristics. Personas with high levels of trust in political and charitable institutions and who are actively engaged in charity and volunteerism prefer donating to human rights organizations that have friendly relationships with their host governments. In contrast, personas with low levels of social trust eschew both human rights organizations *and* any type of NGO with more contentious relationships with host governments. Instead, they prefer donating to NGOs with friendly host-country relationships and working on less contentious issues like emergency response and refugee relief. Personas with higher trust are more comfortable donating to legally besieged organizations. Our findings imply that individual donors with high trust would remain supportive of INGOs when they face criticism and crackdown from foreign governments. On the other hand, government crackdown may further sour INGOs' least trusting donors.

This study contributes both to scholarship and practice in public policy, nonprofits, and philanthropy in three important ways. It helps us understand how the international community and NGOs might respond to the phenomenon of closing civic space. Most NGOs working internationally have previously relied on government and foundation funding. But in the face of decreasing funds from official donors, our results show that INGOs can rely on individual donors—particularly their most trusting donors—when organizations face a hostile environment abroad. These findings lend support to previous literature arguing that trust matters  when individuals donate to organizations acting under great uncertainty or working towards difficult goals [@Wiepking:2010].

Second, knowledge of demographic traits and its impact on funding can help nonprofits craft more effective and targeted marketing strategies [@Rajan:2009, 414]. Our paper shows that in readjusting their fundraising strategies towards individuals, framing, or the process through which actors present information to influence perceptions, is particularly important [@chong2007framing]. Framing affects their willingness to donate. However, existing research is unclear about which frames are effective when making appeals to individual donors considering international philanthropy. Since our study looks at multiple organizational and host country factors, our results can help INGOs frame their appeals. In particular, our results show that INGOs may benefit from publicizing crackdown at least towards their most trusting donors.

Finally, while research in public policy examines giving to NGOs, the discipline of international relations tends to see philanthropy as less relevant [@youde2019]. However, philanthropy—particularly donations received by INGOs—can help us understand changes in the dynamics of global governance. Private individuals exercise power and influence in unique ways that differ from other non-state actors. As @youde2019 [p. 44] points out, studying philanthropy changes our conceptions of individuals being passive recipients to active entities that are not merely objects of governance, and can thus shed light on the changing logic of global governance.

@@

Foreign aid to civil society—and NGOs in particular—has steadily increased over time. Many government and foundation donors perceive NGOs as nimbler and less bureaucratic than governments. In countries with weak institutions and poor governance, where direct aid transfers to governments raise issues of misuse and bureaucratic inefficiency, donors specifically seek out NGOs for projects [@Dietrich2013]. However, NGO-state relationships have also evolved—in the 1950s and 60s, international assistance channeled through NGOs aimed at producing socioeconomic change rather than political progress. But in the aftermath of the dissolution of the Soviet Union, many NGOs had the explicit goal of promoting democracy and bringing down authoritarian governments. Subsequently, many states see NGOs as a threat and seek to repress them.

More than 100 countries have proposed or enacted 244 measures restricting, repressing, or shutting down civil society since 2013 [@icnl:2021]. This has had a threefold impact on INGOs. First, it has led to a reduction in resources and grants available to INGOs. Second, restrictive NGO laws can affect INGO programming by preventing organizations from working in certain (contentious) issue areas. To deal with this, some groups have adapted by changing the nature of their programming to appear less threatening to the repressive countries. Finally, in the worst-case scenario, this has led to a loss in INGOs' access to target countries—often also associated with the use of violence against INGOs and their employees. We focus exclusively on the first dilemma—the reduction in INGO funding and possible responses to it.

@@

INGOs receive funds from a variety of public and private sources. Public funds—or traditional donor aid—channel money from official aid agencies through a variety of bilateral and multilateral institutions. Private sources include foundations, private corporations, and individuals. While INGOs have not been passive in responding to the crackdown on civil society, they face increasing obstacles in funding. Closing civic space has resulted in net losses of income to INGOs, particularly in official bilateral and multilateral aid [@DupuyPrakash:2018]

Legal crackdown has also affected how and where funds are allocated. Following the passage of anti-NGO laws, official donors move funds away from contentious causes such as human rights, media, corruption, advocacy, among others, towards tamer and more regime-compatible causes such as health, education [@ChaudhryHeiss:2018]. The latter trend has also been exacerbated by a preference among donors working in official aid and the democracy promotion establishment to gradually increase funding directed towards tame causes, at the expense of more contentious causes. This is because many tame causes can have easily quantifiable, measurable outputs, which makes it easier to show a program's effectiveness over time [@Bush:2015].

We argue that individual private donors can help fill in an important gap. Individual donors do not use the same performance-based metrics as official donors [@desaikharas]. Rather, individual donors often use heuristics to simply their decision-making, making cursory judgments about an organization's issue area, mission, vision, and values and seek out supplementary information from friends, family, and acquaintances [@ChaudhryHeiss:2021; @TBPrakash:2017; @SzperPrakash:2011]. Due to these differences, individual donors can help address some of the challenges arising from reduced public funding available to INGOs. To be clear, we are not arguing that individual giving would be a perfect substitute for NGOs' net loss of income—rather, private donations can be an important supplement when facing reduced funding. Moreover, private philanthropy can also supplement foundation funding, which has been less adversely affected by hostile legal environments [@NeedlesWolcheckSato:2018].

Research on the determinants of individual-level philanthropy for international causes—particularly towards organizations working in challenging environments—is still in its infancy due to the absence of high-quality data [@greenhill2013]. Moreover, a majority of research on giving to international groups is restricted to elite high net-worth donors who earn more than \$200,000 annually or have more than a \$1 million in assets [@USTrust2014]. We know little about demographic traits and experiences that influence the choice to donate to organizations facing shrinking civic space. Below, we look at key drivers of individual-level philanthropy and highlight the importance of trust when donating during legal crackdowns.

@@

Social capital, particularly associational capital, or "the networks of formal and informal relationships to which people are associated," is an important indicator of giving [@havens2007, 240]. Volunteering exposes individuals to the need for donations, making them feel more compelled to donate [@hossainlamb2012]. Those who participate in a variety of associations are both more likely to donate as well as make larger donations [@hossainlamb2017].

Higher education also influences philanthropic decisions as it fosters prosocial motivations. Education brings people into social networks that entail a higher level of solicitation, which is vital in giving to international causes [@BekkersWiepking:2011]. Education also helps determine the kinds of causes that people donate to globally. @Micklewright:2009 show that higher education matters more for donating to international relief organizations than donating to domestic organizations. A number of NGOs facing crackdowns abroad also work on issues that may be seen as "contentious," including human rights, elections, corruption, advocacy, and media freedom—issues that may have a harder time attracting funds due to their non-essential nature when compared to causes like health, education, sanitation, and social services [@Bush:2015]. However, additional education has been shown to increase abstract thinking [@wiepkingmaas, 1978], which may be necessary to donate to more contentious causes without easily quantifiable goals.

@@

We structure our analysis around broader themes in previous research on individual determinants of philanthropy and we investigate how donors' (1) demographic characteristics, (2) public affairs knowledge and experience, and (3) levels of trust in government, charitable work, and associational life each shape individual donation preferences when simultaneously interacted with different organizational issue areas, relationships with host governments, and funding sources. Our experimental approach provides rich quantitative details of individual and organizational characteristics and reveals unique insights into private donor motivations.

@@

We explore the interaction between individual and organizational characteristics using a conjoint experiment conducted with a nationally representative sample in the U.S. Conjoint analysis allows us to elicit donor preferences without respondents explicitly stating their exact preferences. This kind of experiment is commonly used in marketing to reveal consumer preferences for products and has become increasingly popular in political science and public policy [@KnudsenJohannesson:2019]. Because we are interested in exploring the complexity behind the decision to donate, we apply conjoint methods to the study of nonprofits and philanthropy, where we can study the simultaneous effects of organizational characteristics and individual attributes on individuals' preferences to donate to an INGO.

In our experiment, we presented respondents with four randomly selected international nonprofit organizations with four randomly assigned features.[^appendix-survey] We used organizations that are associated with a range of contentious and noncontentious issues: Amnesty International, Greenpeace, Oxfam, and the International Committee of the Red Cross. Because of their strong name recognition, these organizations (and others like World Vision, the World Wildlife Fund, and Médecins Sans Frontières) are commonly used in experimental surveys on philanthropy [@FaulknerTruongRomaniuk:2015].

@@

```{r organization-attributes-full}
tbl_org_attributes_caption <- "Organization attributes varied in the experiment. Each participant saw twelve sets of four combinations of attributes, with one attribute randomly selected from each column."

@@

After collecting information about demographics and attitudes towards charities, volunteering, and social interests, we presented respondents with twelve sets of four hypothetical organizations and asked them to select the one they would most likely donate to. Each organization was randomly generated with one feature from each of the columns in Table \@ref(tab:organization-attributes-full): issue area, organizational practices, funding sources, and relationship with host government. While not every one of the 288 (4 × 4 × 2 × 3 × 3) possible configurations necessarily reflects reality—for instance, a respondent might have been asked to compare a fictional Amnesty International that focuses on environmental issues and that is funded by private donations with a Greenpeace that focuses on human rights and that is funded by government grants—it is not important that these features align with true organizational characteristics, since the goal of these repeated hypothetical questions is to identify which attributes are the most salient for donors.

In June 2019, we recruited survey respondents through Centiment, a commercial provider of high-quality, non-probability opt-in survey panels. Centiment ensures panel quality by actively recruiting and paying representative samples of the U.S. population. To participate in the study, respondents were first screened based on their charitable activities—those who indicated that they gave to charity once every few years or less were disqualified. After screening, we collected a sample of 1,016 respondents, which is sufficient for model estimation.[^sample-size]

@@

We analyze the effect of different combinations of organizational features and individual respondent attributes with a two-level hierarchical Bayesian multinomial logit model.[^model-details] Simply put, the first level of the model predicts individual donor preferences for various combinations of organizational features (e.g. the combination of human rights issues, financial transparency, government funding, and government crackdown), while the second level predicts the heterogeneity of these fitted preferences as a function of individual donor characteristics (e.g. education, age, and attitudes towards charity, voluntarism, and religion).

@@

We then combine these different scenarios of donor and organization attributes with the coefficients from the hierarchical regression model to calculate each persona's probability of donating to specific combinations of organizational attributes. Any individual characteristics that we do not explicitly vary in Table \@ref(tab:persona-attributes) are held at their means or modal values from the survey sample. Table \@ref(tab:sim-output) provides an example of the simulation output, showing how two different personas are predicted to allocate donations to organizations with various characteristics. The proportion of donations from each persona sums to 100%, as every simulated persona is guaranteed to make a donation.

@@

Simulation methods offer useful analytic benefits. Instead of interpreting marginal coefficient effects from a regression model, we can explore multiple moving parts from the same model simultaneously to understand how individual characteristics interact with each other. This allows us to more fully explore our key research questions and see which kinds of donors are more or less likely to consider specific organizational attributes when deciding to donate, and which donor characteristics are associated with the propensity to donate. Importantly, each persona is an artificial construct. While some survey respondents might match an exact persona profile, many do not. This is standard in simulations as each combination of persona characteristics is plausible in the real world. These simulated findings are more abstract than the results from an ordinary survey sample, but the simulation results allow us to explore the interaction between donor preferences and organization characteristics, including whether they face legal restrictions abroad. The results ultimately show which combination of organizational features each persona is most likely to prefer when choosing to donate.

@@

We examine the results of the simulation in three stages, varying persona (1) demographics, (2) views on politics and public affairs, and (3) social views (i.e. the three columns of Table \@ref(tab:persona-attributes)). Within these three broader categories, we explore the interaction of persona characteristics with different organizational configurations of issue areas, funding sources, and relationships with government (i.e. the three columns of Table \@ref(tab:organization-attributes)).

The four panels of Figure \@ref(fig:income) show the average predicted market share for different organizational issue areas, funding sources, and relationships with host governments across different levels of persona income, with two different cutpoints for high and low income. To ease with modeling and interpretation, we split high- and low-income personas based on 2017 U.S. median income (\$61,372). Due to wide variation in donor incomes in the U.S., we also show the results from a simulation model with a more extreme split, holding annual persona income at \$50,000 and \$100,000+. The difference in the division of high and low income makes no difference in the market shares across issue areas and funding sources, but there are substantial income-based differences across different relationships with host governments.

@@

Next, we next examine how differences in public affairs knowledge, experience, and political ideology influence a persona's propensity to donate to an organization. We vary personas along two dimensions: (1) political ideology, with semi-liberal and semi-conservative political leanings (2 and 5 on a 7-point scale), and (2) knowledge and experience with public affairs and the news, with personas that either follow national and international news and that have traveled abroad or personas that do not follow the news or travel.

We combine these dimensions with an additional set of donor characteristics that vary a persona's level of social trust. Personas with high levels of trust support political institutions and charities, think people should be more charitable, frequently volunteer, donate to charity once a month, have a history of political activism, and are members of an association. Personas with low levels of trust have the opposite characteristics: they do not trust political institutions or charities, rarely donate or volunteer, and are not members of associations.

@@

Personas with high levels of trust, on the other hand, seem to stick with legally besieged organizations and show targeted groups the benefit of the doubt. The global crackdown on NGOs thus sours organizations' least trusting individual donors. Overall, our results support research that donors with high levels of trust are more active in giving than those with low levels of trust [@sargeant2008] and that trust especially matters when individuals donate to organizations working under great uncertainty and towards difficult goals [@Wiepking:2010].

@@

Our conjoint experiment raises two main concerns for inference: (1) prior knowledge and opinions about these organizations could influence respondents' choices, and (2) there are questions about the generalizability of these results outside the American context. We explore both below.

All four organizations from the experiment are well-known, and it is likely that participant responses reflect prior beliefs and news coverage about Amnesty, Oxfam, Greenpeace, and the Red Cross. For instance, in February 2018, news broke out about sexual abuse and misconduct at Oxfam and several other prominent organizations, including Amnesty International and Save the Children. This may have potentially strengthened a narrative that the nonprofit sector suffers from governance failures [@Phillips:2019]. In a competitive NGO environment [@Cooley:2002], maintaining positive perceptions is important as donors who perceive these organizations as untrustworthy may simply take their resources elsewhere. As such, it could be argued that respondents may be less likely to donate to Oxfam or Amnesty because of these scandals.

@@

@@

[^phildef]: We use "philanthropy" to connote voluntary giving to promote the common good. We focus on private individual giving rather than foundation or corporate giving. Some perspectives see philanthropy as a strategic long-term practice addressing the root causes of social issues, and charity as a short-term practice focused on providing immediate relief. But many scholars see philanthropy as essentially synonymous with charitable donations, as short-term giving can also address longer-term or strategic issues [@Castle:2004]. We use the terms interchangeably since individuals could be giving to legally besieged nonprofits for either purpose.
